---
title: "Coursera Data Analytics Capstone Project"
author: "Kacie Morinaga"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---






This R notebook explains my process for the case study for the Coursera Data Analytics Capstone project in Google's Data Analytics Certificate Course.

This case study will be processed using Google's Data Analytics 6 steps: Ask, Prepare, Process, Analyze, Share, and Act.

# Step 1: Ask

In the first step, the purpose of this case study will be explained along with the objectives of the data analysis.

### 1.0 Introduction

Bellabeat is a small, but successful high-tech manufacturer of health-focused products for women. They are looking to expand into the smart device arena to grow their business. As a new junior data analyst you have been asked to examine smart device usage data to gain insight into how consumers are using their smart devices and use these insights to guide marketing initiatives.

### 1.1 Identify the business task

The business task is to identify trends in smart device usage to guide marketing strategy initiatives for Bellabeat

### 1.2 Key stakeholders

-   Urška Sršen: Bellabeat's cofounder and Chief Creative Officer
-   Sando Mur: Mathematician and Bellabeat's cofounder; key member of the Bellabeat executive team
-   Bellabeat marketing analytics team: a team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat's marketing strategy

# Step 2: Prepare

In this step, we identify the data we are using and organize it to make it useful. I decided to use R to perform my data analysis.

### 2.0 Information about the dataset

1.  The dataset is stored online, available [here](https://www.kaggle.com/datasets/arashnic/fitbit) on Kaggle.com, which is an online community platform for data scientists and machine learning enthusiasts.
2.  The dataset is organized into 18 csv files. Generally, it is in long format, but there are wide versions of some of the files. The files I have chosen to use for my analysis are dailyActivity_merged.csv and sleepDay_merged.csv, which are both in long format.
3.  Bias and credibility of the dataset
    -   There are bias and credibility issues with the data because it was collected from 30 FitBit users, which is not reflective of the population of potential Bellabeat customers or the population in general. Additionally it was compiled by a third party so the credibility of the data is questionable.
    -   Does the data ROCCC?
        -   Reliability LOW - The eligibility of the study participants is not specifically established, and the subjects' demographics are unknown. Analysis and subsequent recommendations may not be ideal because Bellabeat is a female focused company and the user group may not be reflective of this population. Additionally, the test group is small at 30 participants, which is again not representative of the general female population.
        -   Original LOW -- the source of the data is from a third party.
        -   Comprehensive MED -- there are several variables collected, including calories burned, daily steps, distance, active minutes, sleep minutes.
        -   Current LOW - the data is not up to date because it was collected from 3.12.2016 to 5.12.2016, over six years ago.
        -   Cited LOW - data collected from a third party.
4.  Licensing, privacy, security, and accessibility - The dataset is public and the creator waived all rights to their work worldwide under copyright law. Therefore, everyone is allowed to copy, modify, distribute, and perform the work.
5.  Data integrity - Because the data is from a third-party, its accuracy cannot be guaranteed. Upon initial glimpse, it is difficult to determine whether or not it is complete, but that can be determined upon further examination. The quality of the data is on the lower end, as determined by looking at the ROCCC scores.
6.  The data can help answer our business task because it gives us two months information of activity tracking using a smart device from 30 users. By analyzing the data, we may be able to find activity and usage trends to improve appeal of these smart devices to potential customers.

### 2.1 Key tasks

#### 2.1.1 Download the data and store it appropriately

I downloaded the dataset from [here](https://www.kaggle.com/datasets/arashnic/fitbit) on Kaggle.com, and created a folder on my computer to keep the raw data files, labeling each appropriately.

#### 2.1.2 Identify how it is organized

I decided to use R to perform my data analysis. In order to identify how it is organized, I needed to import it to R. First, I have to set up my environment.

##### Setting up my environment

I've chosen to load the entire tidyverse. It generally contains all the packages I will need for analysis. If I come across another package I need in the future, I can load it at that time.

```{r Loading packages}
library(tidyverse)
```

##### Importing our data

I've downloaded the dataset to my computer as csv files from kaggle.com. I can import it into R using readr, which was in the tidyverse package. I've chosen to use the dailyActivity_merged and sleepDay_merged dataframes.

```{r Using readr to import dataset}
daily_activity <-read_csv("/Users/kaciemorinaga/Downloads/dailyActivity_merged.csv")
sleep_day <- read_csv("/Users/kaciemorinaga/Downloads/sleepDay_merged.csv")
```

As I imported the data, I changed the names of the dataframes to a consistent format, using only lowercase letters and underscores in place of spaces.

It is revealed that there are 940 rows with 15 columns in the daily_activity dataframe. There are 413 rows with 5 columns in the sleep_day dataframe. This is important to note because it seems that there is more than double the data available for daily activity vs. sleeping.

### 2.2 Previewing the data

##### 2.2.1 The glimpse function

The glimpse function allows you to see the titles of all of the columns of your dataframes and lists the first few variables of each.

```{r Using glimpse}
glimpse(daily_activity)
glimpse(sleep_day)
```

Using glimpse on the dataframes reveals that both have an Id column and followed by a date column, although in daily_activity it is "ActivityDate" and in sleep_day it is "SleepDay". Additionally, the SleepDay data also has a time component.

##### 2.2.2 The str function

The str function shows the structure of the dataset. We can use it to examine both dataframes.

```{r Using str}
str(daily_activity)
str(sleep_day)
```

Of note is that the date in both dataframes are formatted as characters, rather than dates. Additionally the id is formatted as a number, rather than a character. These will be converted later in the cleaning process.

##### 2.2.3 Looking for null values

Knowing if null values are present and how many there are will tell me if there is data missing. I decided to look at the dataframes separately so I would know if there was data missing in the raw form, and which dataframe is missing information.

```{r Looking for null values in the two dataframes}
sum(is.na(daily_activity))
sum(is.na(sleep_day))
```

Now we know that both tables have no null values when examined on their own.

### 2.3 Sorting and filtering the data

So far, the data seems to be in a decent order in both dataframes. It is sorted first by id, then by date and then shows all the variables collected for the listed subject on the specified date. As mentioned previously there are some formatting changes to be made, but they can be done during the cleaning process.

### 2.4 Renaming the dataframe columns

Following the standardized format used to rename the dataframes, I have also decided to format all the columns using all lowercase letters and underscores to mark the spaces between words. However, I will have to install and load the janitor package to use the clean_names function.

```{r Loading janitor package}
library(janitor)
```

Once I have installed the janitor package I can use the clean_names function to rename the columns.

```{r Cleaning the column names}
daily_activity <- clean_names(daily_activity)
sleep_day <- clean_names(sleep_day)
```

Running glimpse again will show the changes.

```{r View changes made with renaming}
glimpse(daily_activity)
glimpse(sleep_day)
```

### 2.5 Converting activity_date and sleep_day from character to date format

Now it's time to change the activity_date column from character format to date format. I've decided to use the as.Date function to convert the data into the correct format.

```{r Formatting the activity_date column from character to date}
daily_activity$activity_date <- as.Date(daily_activity$activity_date, format = "%m/%d/%Y")
```

The sleep_day column, however, has an additional time component which will need to be removed before converting to date format. This can be taken care of with the gsub function, nested within the as.Date function.

```{r Formatting the sleep_day column from character to date}
sleep_day$sleep_day <- as.Date(gsub(" 12:00:00 AM.*", "", sleep_day$sleep_day), format = "%m/%d/%Y")
```

### 2.6 Converting user ids from number to string characters

Because the user ids are not numerical values, I decided to convert them to string characters for ease with later analysis.

```{r Converting user ids to string characters}
sleep_day$id<-as.character(sleep_day$id)
daily_activity$id<-as.character(daily_activity$id)
```

### 2.7 Combining the dataframes

Now that we have two dataframes with id and date columns, we can combine them. I decided to use a left join to combine the dataframes, retaining all the records from the daily_activity dataframe because it has more rows and therefore more data entries. These will be combined with the matching sleep_day data, if available. Keeping the daily_activity information without matching sleep_day data will provide us additional insight because it demonstrates that the device was used during the day, but not at night.

```{r Combining the dataframes}
lj_combined <- left_join(daily_activity, sleep_day, by=c("id"="id", "activity_date"="sleep_day"))
```

Let's take a look at the new dataframe.

```{r Looking at the combined dataframe}
glimpse(lj_combined)
```

It looks good so we can proceed onto the next step of our analysis.

# Step 3: Process

In this step, the data will be further cleaned and transformed in preparation for analyzation.

### 3.0 Create new day_of_week column

In general, most people tend to work Monday through Friday, which can affect their activity levels. Therefore, it may be useful to look at weekday vs. weekend data to see if the days of the week will affect smart device use.

#### 3.0.1 Loading lubridate package

In order to use the function I want, I need to install and load the lubridate package.

```{r Loading the lubridate package}
library(lubridate)
```

#### 3.0.2 Creating the day_of_week column

The wday function will allow me to create a new column with the day of the week to match each date. 

```{r Creating new day_of_week column}
lj_combined$day_of_week = c(wday(lj_combined$activity_date, label=TRUE))
```

#### 3.0.3 Reordering the dataframe to put day_of_week after activity_date

Because day_of_week is determined from the date, it is best to have these columns next to each other.

```{r Reordering the dataframe}
lj_combined <- lj_combined[, c(1,2,19,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18)]
```

### 3.1 Converting minutes to hours

I am using the mutate function to convert minutes to hours because we typically think of time in hours when the number of minutes is in the hundreds.

```{r Converting minutes to hours}
lj_combined <- mutate(lj_combined, very_active_hours = very_active_minutes/60, fairly_active_hours = fairly_active_minutes/60, lightly_active_hours = lightly_active_minutes/60, sedentary_hours = sedentary_minutes/60, total_hours_asleep = total_minutes_asleep/60, total_time_in_bed = total_time_in_bed/60)
```

#### 3.1.1 Deleting the minutes column and reordering the dataframe

Now that I have the new hours columns, I can delete the minutes columns and reorder the remaining columns.

```{r Deleting minutes columns and reordering the dataframe}
lj_combined <- lj_combined[, c(1,2,3,4,5,6,7,8,9,10,11,20,21,22,23,16,17,24,19)]
```

Now that the data has been cleaned and transformed let's take a look at the finalized dataframe before we start our analysis.

```{r Looking at the finalized dataframe}
glimpse(lj_combined)
```

The dataframe looks good so we can proceed to the next step.

# Step 4: Analyze

In this step, we will look at some basic statistics from the dataset and other numbers to find trends and relationships in the data. I will also be creating a couple simple tables to get a different perspective of the same information.

### 4.0 Using the summary function to look at generic statistics of the dataset

The summary function will give me a better idea of the distribution of the variables in my dataset. I picked a few of the variables that I felt might be the most informative.

```{r Summary of the dataframe}
lj_combined %>% 
  select(day_of_week, total_steps, tracker_distance, logged_activities_distance, very_active_hours, fairly_active_hours, lightly_active_hours, sedentary_hours, calories, total_sleep_records, total_hours_asleep, total_time_in_bed) %>% 
  summary()
```

There are a few interesting insights to gain from this summary of the dataset.

1.  The mean total_steps was 7,652 which is below the [CDC](https://www.cdc.gov/diabetes/prevention/pdf/postcurriculum_session8.pdf) recommended 10,000 steps per day.
2.  The mean logged_activities_distance was 0.110 versus the mean tracker_distance which was 5.489, demonstrating that users are mostly passively using the devices to log their activities.
3.  According to the [2015-2020 Dietary Guidelines for Americans](https://health.gov/sites/default/files/2019-09/2015-2020_Dietary_Guidelines.pdf), women are likely to need to intake an average of 2,000 calories per day, men 2,500 calories per day. The mean calories burned for this dataset is 2,308, nearly in the middle of the recommended intake of calories for both genders. Assuming users are consuming the recommended calories, it would seem they are not using the devices to lose weight, as each pound of fat accounts for 3,500 calories.
4.  The top three days of the week with the most records are in order, Tuesday, Wednesday, and Thursday, demonstrating that device use is most active mid-week.
5.  There are 530 NA's for the sleep data. Because there are 943 total records, that means the devices are being used less than half the time at night.

### 4.1 Count unique users in the dataset

Per the dataset source, there is data from 30 eligible users, but it is important to confirm the number of users.

```{r Count unique users}
n_distinct(lj_combined$id)
```

By running this chunk, it is revealed that there are actually 33 unique users, not 30 as stated by the dataset author. This will be important to remember in later analysis, when we look at averages among the users.

I also thought it was important to count the number of unique users with sleep data. When we use the sleep data from the lj_combined dataframe, this will be the number of users because of the daily activity records that did not have matching sleep data.

```{r Count unique sleep data users}
lj_combined %>% 
  filter(total_sleep_records!="NA") %>% 
  summarize(count=n_distinct(id))
```

This reveals we have 24 users with complete datasets.

### 4.2 Daily device usage - full vs. partial data

I also thought it was important to look at how often users were actually using their devices throughout the whole day. In order to make this determination, I chose to look at the number of active_hours plus the number of sedentary_hours that added up to 24 as well as those that did not add up to 24 because there are 24 hours in a day.

```{r Full vs. partial data per day}
sum(lj_combined$very_active_hours + lj_combined$fairly_active_hours + lj_combined$lightly_active_hours + lj_combined$sedentary_hours == 24)
sum(lj_combined$very_active_hours + lj_combined$fairly_active_hours + lj_combined$lightly_active_hours + lj_combined$sedentary_hours != 24)
```

This reveals that 478 out of 943 records have data for the full day and 465 records have data for part of the day. Therefore, about half the time users are not using the device all day. This is likely related to the fact that there are about 530 records that are missing sleep data and users are not consistently using the devices to track sleep.

It is also interesting to note that there are 530 entries that are missing sleep data, but only 465 entries with partial data. This demonstrates that the devices may not always be properly recording sleep data.

### 4.3 Averages by id

While overall averages of our user group records are useful to know, the typical activity level and sleep amounts of each individual can vary greatly and combine to make an average that looks unremarkable. The summary function also reinforces this idea, revealing that the minimum calorie record was 0 and the maximum record was 4,900. Because of this vast difference in records, I thought it would be important to look at the averages of multiple variables by id. I decided to create a new dataframe for this information, which we can later use during visualization.

```{r Creating a new table for averages per id}
avgs_per_id<-lj_combined %>% 
  group_by(id) %>% 
  summarize(avg_total_steps=mean(total_steps), avg_tracker_distance=mean(tracker_distance), avg_logged_activities_distance=mean(logged_activities_distance), avg_active_hours=mean(very_active_hours+lightly_active_hours+lightly_active_hours), avg_sedentary_hours=mean(sedentary_hours), avg_calories=mean(calories), avg_total_hours_asleep=mean(total_hours_asleep, na.rm=TRUE), avg_total_time_in_bed=mean(total_time_in_bed, na.rm=TRUE))
```

For this new table, I combined very_active_hours, fairly_active_hours, and lightly_active_hours to create a new avg_active_hours for each id because I felt that it was more informative to look at overall active time rather than having it split into levels. 

Here's a look at the first few entries of the new dataframe.

```{r Looking at avgs_per_id dataframe}
head(avgs_per_id)
```

### 4.4 Average calories by day of week

Because of my theory about weekday versus weekend calorie differences, I felt it would be important to examine average calories per day of week as well. Therefore I created another dataframe with this information, also for use during analysis.

```{r Creating a new table for average calories per day of week}
avg_calories_per_day_of_week<-lj_combined %>% 
  group_by(day_of_week) %>% 
  summarize_at(vars(calories), list(avg_calories=mean))
```

And here's a look at the new dataframe.

```{r Looking at calories_per_day_of_week dataframe}
head(avg_calories_per_day_of_week, n=7)
```

# Step 5: Share

In this step, visualizations will be created to demonstrate insights from the dataset, which will help us to make recommendations to the company.

### 5.0 Count by day of the week

The typical workweek is Monday through Friday and most people are spending eight hours a day at work, leaving less free time to exercise on those days. Therefore, I thought it would be important to examine device usage to see if there was a difference between weekday vs. weekend usage.

```{r Records by day of the week}
ggplot(lj_combined, mapping=aes(x=day_of_week, fill=..count..))+
  geom_bar(width=0.7, stat="count")+
  scale_fill_gradient(low="#132B43", high="#56B1F7")+
  geom_text(stat="count", aes(label=..count..), vjust=-.25)+
  labs(title="Records per Day of Week", x="Day of the Week", y="Count", fill="Count")
```

This bar graph demonstrates that device usage is highest on Tuesday, Wednesday, and Thursday, in that order. While there is no certainty as to why usage is highest mid-week, it may indicate that devices are being used to look at activity levels on workdays so users can be aware to try to increase activity if they are falling short of their goals, which is more likely when a large portion of their day is devoted to work.

### 5.1 Average calories per day of week

This prompts me to look at the average calories per day of week, on the same assumption of a typical Monday through Friday workweek with weekends on Saturdays and Sundays.

```{r Visualization for average calories per day}
ggplot(avg_calories_per_day_of_week, mapping=aes(x=day_of_week, y=avg_calories, fill=avg_calories))+
  geom_col(width=0.7)+
  scale_fill_gradient(low="#132B43", high="#56B1F7")+
  geom_text(aes(label=sprintf("%.1f", avg_calories), vjust=-.25))+
  labs(title="Average Calories per Day", x="Day of the Week", y="Calories", fill="Calories")
```

This visualization shows that Tuesday, Wednesday, and Thursday are *not* the top three days of the week for calorie burn. There is relatively little variation in the average amount of calories burned per day of week - the difference between the minimum and maximum is 161.3 calories. The fact that there is little variation in the average calories burned per day of week may indicate that users are balancing their activity levels as there is no significant differentiation in calorie burn during weekdays vs. weekends. 

### 5.2 Average calories per id

As mentioned previously, the typical activity level of each individual can vary greatly and combine to make an average that looks unremarkable. Therefore, I previously created a table with average calories per user id. Now we can look at the data from that table and compare it to recommended daily calorie intakes by gender.

#### 5.2.1 Average calories vs. recommended male intake

```{r Visualization for average calories per id vs. recommended male intake}
ggplot(avgs_per_id, mapping=aes(x=id, y=avg_calories, fill=avg_calories>2500))+
  geom_col(show.legend=FALSE)+
  scale_fill_manual(values=c("darkgray", "#00BA38"))+
  geom_hline(aes(yintercept=2500, color = "blue"), linetype=1, color="blue", size=1)+
  geom_hline(aes(yintercept=2308, color="purple"), linetype=2, color="purple", size=.5)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="Average Calories per User vs. Recommended Intake for Males", x="Users", y="Calories")
```

I added in a couple of horizontal lines to show how the user group compared with the recommended daily calorie intake for males (solid blue) and the overall average calories burned for the group (dashed purple). Looking at the chart, most users are clearly falling short of burning the recommended calorie intake for males. Here's the actual number of users in the group that did not burn an average above the recommended intake for males.

```{r N for calorie averages below recommended intake for males}
nrow(subset(avgs_per_id, avg_calories<2500))
```

There are 21 out of 33 in the user group who are averaging less than 2500 calories burned according to data collected by the smart devices.

#### 5.2.2 Average calories vs. recommended female intake

Let's take a look at how the user group compared with the recommended daily calorie intake for females, which is 2000 calories.

```{r Visualization for average calories per id vs. recommended female intake}
ggplot(avgs_per_id, mapping=aes(x=id, y=avg_calories, fill=avg_calories>2000))+
  geom_col(show.legend=FALSE)+
  scale_fill_manual(values=c("darkgray", "#00BA38"))+
  geom_hline(aes(yintercept=2000, color="red"), linetype=1, color="red", size=1)+
  geom_hline(aes(yintercept=2308, color="purple"), linetype=2, color="purple", size=.5)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title="Average Calories per User vs. Recommended Intake for Females", x="Users", y="Calories")
```

This chart is similar to the previous one, but the recommended daily calorie intake for females is represented by the solid red line. The purple dashed line again represents the overall group average of calories burned. More users are burning a daily average over the recommended calorie intake for females compared to the recommended intake for males. Let's see how this number compares.

```{r N for calorie averages below recommended intake for females}
nrow(subset(avgs_per_id, avg_calories<2000))
```

#### 5.2.3 Insights:

1.  21 out of 33, 63.6% of users are burning less than the recommended calorie intake for males.
2.  13 out of 33, 39.3% of users are burning less than the recommended calorie intake for females.

If users are consuming around the recommended number of calories, most are not likely using the devices to lose weight. A majority of them are burning less than the recommended intake for males, and while not a majority, there is still a good percentage that is not burning the recommended calorie intake for females.

### 5.3 Examining logged activities distance and tracked distance

There is an option for users to log activities, but the device also passively tracks distance for users. Looking at the data for these two variables can help determine how users tend to use the devices.

```{r Visualization for logged and tracked activities distance}
ggplot(lj_combined, mapping=aes(x=tracker_distance, y=logged_activities_distance, color=id))+
  geom_point(show.legend=FALSE)+
  labs(title="Logged vs. Tracked Activity Distance", x="Tracker Distance", y="Logged Activities Distance")
```

This chart demonstrates that users are generally passive in regards to tracking distance. The majority of the time users have not logged in any distances, however the smart devices have been passively keeping track of the users distance.

Looking at this chart, it would be interesting to know how many "0" entries there are for logged distance because they are all muddled at the bottom of the chart.

```{r Determining how many records have 0 for logged distance}
nrow(subset(lj_combined, logged_activities_distance == 0))
```

Out of 943 entries, there are 910 times that the logged_activities_distance was 0. This means users are logging activity very infrequently - in fact, only about 3.5% of the records have logged activity. It may be useful to market to users that the devices track activities passively as it seems to be the preferred means of this user group.

It would also be interesting to see how many users out of the group are actually logging activities.

```{r Number of users tracking activities}
nrow(subset(avgs_per_id, avg_logged_activities_distance>0))
```

Four users out of the 33 are actually logging activities, which equates to 12.1% of the user group. While this is not a large percentage, it shouldn't be ignored because there are a few individuals from this small user group that use the smart devices to log activities. There is a potential to lose over 10% of customers if they are unaware that the devices perform this functionality.

### 5.4 Tracker Distance and Active Hours

Another way the smart device logs activity is by time. I previously converted the active minutes into hours to make the measurement more relatable as people typically divide their day into hours, not minutes. We already know that users did not frequently log their activies, so we'll examine tracker distance and active hours.

```{r}
ggplot(lj_combined, mapping=aes(x=(very_active_hours+fairly_active_hours+lightly_active_hours), y=tracker_distance))+
  geom_point()+
  labs(title="Tracker Distance vs. Active Hours", y="Tracker Distance", x="Active Hours" )
```

As expected, tracker distance tends to increase as active hours also increases. There may be some inconsistency in recording active time as there are some outliers where the user has distance tracked, but no active time. 

### 5.5 Average hours asleep per user

The smart devices also have the capability of tracking time asleep, which is another important marker of health behavior. According to the [CDC](https://www.cdc.gov/sleep/features/getting-enough-sleep.html), not getting enough sleep regularly can lead to an increased risk of obesity, type 2 diabetes, high blood pressure, heart disease and stroke, poor mental health, and early death. Let's look at the sleep data of our user group.

```{r Visualization for average hours asleep per user}
avgs_per_id %>% 
  drop_na(avg_total_hours_asleep) %>% 
  ggplot(mapping=aes(x=id, y=avg_total_hours_asleep, fill=avg_total_hours_asleep>7))+
  geom_col(show.legend=FALSE)+
  scale_fill_manual(values=c("red", "darkgray"))+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  geom_hline(yintercept=7, color="#00B159", size=1)+
  labs(title="Average Hours Asleep per User", x="Users", y="Average Hours Asleep")
```

For adults over 18, the CDC recommends seven or more hours of sleep per night, represented by the green line in this visualization. The red bars highlight the users that are falling short of this recommendation. Let's determine the number of users not getting enough sleep.

```{r N for users getting less than 7 hours sleep per night on average}
nrow(subset(avgs_per_id, avg_total_hours_asleep<7))
```

Exactly half (12/24) of the users in the group with available sleep data are not getting the recommended amount of sleep per night on average. Bellabeat can use this information to encourage use of their devices to track sleep and improve on this particular health behavior, as 50% of the sample group is falling short of the recommended amount.

According to [CDC 2020 data](https://www.cdc.gov/sleep/data-and-statistics/adults.html), the crude prevalence of adults who reported short sleep duration, defined as less than seven hours per night, is 32.8%. This number is self-reported as participants were asked how many hours of sleep they got in a 24 hour period. The smart device user group has a higher percentage of participants averaging less than seven hours of sleep a night, but their data is likely to be more accurate because it is recorded in real time by the smart devices. 

# Step 6: Act

In this step, I will share final conclusions from my analysis and make suggestions on ways the team and business can apply these insights.

### Key Conclusions

1.  The days with the most records are mid-week, generally when most people are working. Additionally, users tend not to log their activities, preferring to have the devices track activities passively. Using both pieces of information, Bellabeat can market smart devices to busy working women. **Bellabeat can emphasize the message that just like these hard-working women, their smart devices are also working hard, but the devices are working quietly in the background, tracking their activity while allowing the women to concentrate on more important things.**
2.  While the majority of users did not log activities using the devices, there is a significant percentage of the group, four out of 33, or 12.1%, that did take advantage of the logging capability of the smart devices. **While acknowledging that Bellabeat tracks activities passively, it should also be mentioned that the devices are flexible and have the ability to log activities if desired.**
3.  The smart devices track calories burned daily and this user group is diverse in this regard. The lowest average calorie burn is about 1500 and the highest nearly 3500. The diversity of the group can be emphasized as a strength of the smart devices. **Whether users are actively working to lose weight or they are just interested in tracking calories burned with normal daily activities, Bellabeat smart devices are made for all types of women.**
4.  About half of the records are lacking data for the entire day, likely because users are not tracking their sleep with the smart devices. On top of this, lack of regular sleep can contribute to the development of a multitude of diseases and over one third of the American adults are falling short of the recommended seven hours per night. The smart devices have the advantage of accurately tracking users' total hours asleep, taking away the guessing game trying to figure out actual sleep and wake times. **Bellabeat can highlight the sleep tracking capabilities of their smart device, helping users to accurately determine their nightly averages. Knowing this information, users can attempt adjust their sleep behaviors to help minimize the risk of developing chronic health diseases and conditions.**

Unaddressed concerns

1. Knowing the demographics of the FitBit user group would be helpful to determine if the user group is reflective of the Bellabeat population. If the groups are vastly different, the insights passed along to the marketing group are less likely to be effective. 
2. The Bellabeat smart devices may offer different features compared to the FitBit devices. It would be useful to know what similar and different features the Bellabeat smart devices offer to highlight and seek case studies with these features to provide more specific insights. 